---
title: "Products"
output: html_notebook
---



```{r}
#read dataset
IFraud = read.csv("Fraud.csv")
head(IFraud)

```

Initial exploration
```{r}
#Initial structure of the dataset
str(IFraud)
```


```{r}
#drop columns that are not necessary
  IFraud = IFraud[,-c(2,8,17:18)]
```


```{r}
  #change the factor variables to factor
   factors <- c(27,28,15,16,14)
   IFraud[,factors] <- lapply(IFraud[,factors], factor)
   #new structure
    str(IFraud)
```


```{r}
#missing values
library(VIM)
sum(is.na(IFraud)) #11 missing values
```


```{r}
#zeroes present in age, days of week claimed, and month claimed
  IFraud$Age[IFraud$Age == 0] = NA
  IFraud$DayOfWeekClaimed[IFraud$DayOfWeekClaimed == "0"] = NA
  IFraud$MonthClaimed[IFraud$MonthClaimed == "0"] = NA
  
  #drop levels ("0" level) that did not occur
   IFraud$DayOfWeekClaimed = factor(IFraud$DayOfWeekClaimed)
   IFraud$MonthClaimed = factor(IFraud$MonthClaimed)
  
  #Missing values now
  sum(is.na(IFraud)) #266 missing values
  
```


```{r}
#visualize where they are
  aggr(IFraud, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(hour), cex.axis=.7, gap=3,   ylab=c("Histogram of missing data","Pattern")) #missing values in age and Driver rating day of week claimed and month claimed
#remove missing values since there are enough observations 
  Fraud = na.omit(IFraud)
```


```{r}
#lets explore the numeric columns Age and claim size
  Numerics = Fraud[,c(9,30)]
  #summary
  summary(Numerics)
```


```{r}
#standard deviation
  lapply(Numerics, sd)
```


```{r}
#outliers
 # library(dplyr)
  #library(tidyr)
  #library(purrr)
#visualize presence of outliers
  boxplot(Numerics)$out  #outliers present in claimsize
  
  #replace outliers with NA
  replaceoutliers <- function(df){
           df %>%          
           map_if(is.numeric, ~ replace(.x, .x %in% boxplot.stats(.x)$out, NA)) %>%
           bind_cols 
      }
  
  #replace outliers in fraud with NA
  Fraud = replaceoutliers(Fraud)
    #remove outliers(NA)
    Fraud = na.omit(Fraud)
```


```{r}
#Visualize the columns
    library(ggplot2)
    library(reshape2)
    ggplot(data = melt(Numerics) , aes(x=value)) +
      geom_histogram() + 
      facet_wrap(~variable, scales = "free")
```


```{r}
#Investigate correlation between variables
    library(GGally)
      ggcorr(Numerics, label = TRUE, label_round = 2, label_size = 3.5, size = 2, hjust = .85) +
      ggtitle("Correlation Heatmap") +
      theme(plot.title = element_text(hjust = 0.5)) #no high correlation
      
```
Question 1
```{r}
 #is there is a significant relationship between age and claimsize (by analyzing the variance)
      summary(aov(Fraud$Age~Fraud$ClaimSize)) #No there isnt
      #visualize the relationship  
      plot(aov(Fraud$Age~Fraud$ClaimSize))
```
Question 2 - what is the average claimsize based on 
*accident area
*month claimed
*sex
*year
```{r}
#Accident Area
aggregate(Fraud[, 30], list(Fraud$AccidentArea), mean)
#sex
aggregate(Fraud[, 30], list(Fraud$Sex), mean)
#Year
aggregate(Fraud[, 30], list(Fraud$Year), mean)

```
Question 3 - what is the age distribution based on driver rating
```{r}
Q3 <- data.frame(Fraud$Age,Fraud$DriverRating)
colnames(Q3) = c("Age","Rating")

## plot histogram
Q3_hist <- ggplot(Q3, aes(x= Rating, fill = Rating)) + 
    geom_bar(position = "dodge", stat = "count")  +
    ggtitle("Driver ratings") +
    scale_fill_manual(values = c("#009E73", "#D55E00","#F0E442","#999999")) + 
    ylab("count") + 
    xlab("People")
library(dplyr)
Q3_ <- group_by(Q3[, c("Age", "Rating")], Rating, Age) %>% summarize(n_Age = n())

Q3_graph <- ggplot(Q3_, aes(x= Age, y = n_Age, color = Rating)) + 
    geom_line(size = 1.5)  +
    ggtitle("What is the age distribution ?") +
    scale_color_manual(values = c("#009E73", "#D55E00","#F0E442","#999999")) + 
    ylab("number") + 
    xlab("age") +
    theme(legend.position="none")

library(gridExtra)
grid.arrange(Q3_hist, Q3_graph, ncol=2)
```
Question 4 - What category of people (based on age) have the highest claim size
```{r}
# Get some plots of the data set
x = Fraud
#categorize claimsize
x$ClaimSize_cut = cut(as.numeric(as.character(x$ClaimSize)), 
                              c((0:5)*20000), right = FALSE, 
                              labels = c("0-19K", "20K-39K", "40K-59K", 
                                         "60K-79K", "80K+"))


ggplot(data=x, aes(ClaimSize_cut,VehiclePrice)) +
  geom_bar(stat = "identity", aes(fill = VehiclePrice), position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs( title = "VehiclePrice Vs ClaimSize", x = "ClaimSize($)", 
        y = "Vehicle Price ($)")


```
Question 5 - price of cars vs claim size
```{r}

#categorize age
x$Age_cut = cut(as.numeric(as.character(x$Age)), c(seq(10,70,20)), right = FALSE)

x = na.omit(x)
ggplot(data=x, aes(ClaimSize_cut, Age_cut), na.rm = TRUE) +
  geom_bar(stat = "identity", aes(fill = Age_cut), position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs( title = "Age and ClaimSize Categories", x = "ClaimSize Categories ($)", 
        y = "Age Categories (Years)")

```
Question 6 - What sex have the highest claim size
```{r}
ggplot(data=x, aes(ClaimSize_cut,Sex)) +
  geom_bar(stat = "identity", aes(fill = Sex), position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs( title = "Sex vs ClaimSize", x = "ClaimSize($)", 
        y = "sex")
```
Question 7 - Age distribution based fraud found
```{r}
# First, given a list for converting the labels 
default_names = list('0' ="NoFraud", '1'= "Fraud")

# Then, define a labeller to convert labels
d_labeller = function(variable,value){
  return(default_names[value])
}

ggplot(data=x, aes(x=Age)) + 
  geom_histogram(binwidth=.5, colour="black", fill="white") + 
  facet_grid(FraudFound_P ~.,labeller = d_labeller) +
  geom_vline(data=x, aes(xintercept=mean(Age, na.rm=T)), 
             linetype="dashed", size=1, colour="red") +
  labs(title = "Histogram of Age and FraudFound_P", x = "Age (Years)",
       y = "FraudFound_P")

#The bar charts shows there is higher percentage of fraud for people between age 16 and age 40. 

```
Question 8 - when can we say insurance fraud can occur ?
```{r}
ggplot(data=x, aes(x=ClaimSize, colour=FraudFound_P)) +
  stat_density(geom="line",position="identity") + 
  stat_density(geom="line", aes(color = "FraudFound_P")) +
  labs(title = "Density of ClaimSize and Insurance Fraud", 
       x = "ClaimSize($)", y = "Density") + 
  scale_colour_discrete(name="Fraud", breaks=c("0", "1", "FraudFound_P"),
                         labels=c("NoFraud", "Fraud", "All (Yes and No)"))
## Comment: Light blue line, which represents the density of Fraud has a high peak at claim size about 5000$. It tells us that frauds are liable to occur when the claim size is at that level.
```
Question 9 - Marital status of Fraudsters (yearly)
```{r}
# Grouped Bar Plot
ggplot(x, aes(fill=MaritalStatus, y=ClaimSize, x=Year, main = "Pecentage accuracy")) + 
    geom_bar(position="dodge", stat="identity") +
  facet_wrap(~FraudFound_P)
```
Question 10 - Past number of claims
```{r}
PNCtable = as.data.frame(table(x$PastNumberOfClaims))
PNCtable$Prob = PNCtable$Freq / sum(PNCtable$Freq)
colnames(PNCtable) = c("PNC", "Freq", "Prob" )
ggplot(PNCtable, aes(x="", y=Prob, fill=PNC)) +
  geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + 
  labs(title = "Pie Chart of PNC", y = "Probability")
## Comment: more than half of the people have made insurance claims previously
```
Modelling -  [logistic regression and randomforest] to predict the claim size of "Fraud" claims and "NoFraud" claims based on sex, marital status, age, pnc, vehicle price, accident area

splitting and subsetting the data set dataset
```{r}
#Nofraud dataset
library(caret)
NoY = Fraud[Fraud$FraudFound_P == 0,]
#scale claimsize
NoY$ClaimSize = scale(NoY$ClaimSize)
i <- createDataPartition(NoY$ClaimSize, p = 3/4,list = FALSE)
train <- NoY[i,]
test <- NoY[-i,]
#Fraud dataset
YesY = Fraud[Fraud$FraudFound_P == 1,]
YesY$ClaimSize = scale(YesY$ClaimSize)
j <- createDataPartition(YesY$ClaimSize, p = 3/4,list = FALSE)
Ytrain <- YesY[j,]
Ytest <- YesY[-j,]
#All
Fraud$claimSizeS = scale(Fraud$ClaimSize)
k <- createDataPartition(Fraud$claimSizeS, p = 3/4,list = FALSE)
Atrain <- Fraud[k,]
Atest <- Fraud[-k,]
#set cross validation and evaluation metric
control <- trainControl(method='cv', number=4)
metric <- 'RMSE'



```
Logistic regression
```{r}
#logistic regression on Nofraud
Nofraud.logit <- train(ClaimSize~Sex+Age+MaritalStatus+VehiclePrice+AccidentArea, data=train, method='glm',
                   trControl=control, metric=metric)
        #results on train set
        View(Nofraud.logit$results)
        #Function to calculate RMSE and mae on test set
        calc_rmse = function(actual, predicted) {
        sqrt(mean((actual - predicted) ^ 2))}
        calc_mae = function(x,y){
          mean(x-y)}
        #rmse
        calc_rmse(actual = test$ClaimSize,
          predicted = predict(Nofraud.logit, test))
        #mae
        calc_mae(x = test$ClaimSize,
                 y = predict(Nofraud.logit,test))
#logistic regression on fraud
fraud.logit <- train(ClaimSize~Sex+Age+MaritalStatus+VehiclePrice+AccidentArea, data=Ytrain, method='glm',
                   trControl=control, metric=metric)
        #results on train set
        View(fraud.logit$results)
        #RMSE
        calc_rmse(actual = Ytest$ClaimSize,
          predicted = predict(fraud.logit, Ytest))
        #MAE
        calc_mae(x = Ytest$ClaimSize,
                 y = predict(fraud.logit,Ytest))
#logistic regression All subsets
Afraud.logit <- train(claimSizeS~Sex+Age+MaritalStatus+VehiclePrice+AccidentArea, data=Atrain, method='glm',
                   trControl=control, metric=metric)
        
        #results on train set
        View(Afraud.logit$results)
        
        #RMSE
        calc_rmse(actual = Atest$claimSizeS,
          predicted = predict(Afraud.logit, Atest))
        #MAE
        calc_mae(x = Atest$claimSizeS,
                 y = predict(Afraud.logit,Atest))

```
RandomForest
```{r}
#Random forest on Nofraud
Nofraud.rf <- train(ClaimSize~Sex+Age+MaritalStatus+VehiclePrice+AccidentArea, data=train, method='ranger',
                   trControl=control, metric=metric)
        
        #results on train set
        View(Nofraud.rf$results)
        
        #rmse
        calc_rmse(actual = test$ClaimSize,
          predicted = predict(Nofraud.rf, test))
        
        #mae
        calc_mae(x = test$ClaimSize,
                 y = predict(Nofraud.rf,test))
        

#RandomForest on fraud
fraud.rf <- train(ClaimSize~Sex+Age+MaritalStatus+VehiclePrice+AccidentArea, data=Ytrain, method='ranger',
                   trControl=control, metric=metric)
        
        #results on train set
        View(fraud.rf$results)
        
        #RMSE
        calc_rmse(actual = Ytest$ClaimSize,
          predicted = predict(fraud.rf, Ytest))
        #MAE
        calc_mae(x = Ytest$ClaimSize,
                 y = predict(fraud.rf,Ytest))
        
#Random Forest on all subsets
Afraud.rf <- train(claimSizeS~Sex+Age+MaritalStatus+VehiclePrice+AccidentArea, data=Atrain, method='ranger',
                   trControl=control, metric=metric)
        
        #results on train set
        View(Afraud.rf$results)
        
        #RMSE
        calc_rmse(actual = Atest$claimSizeS,
          predicted = predict(Afraud.rf, Atest))
        #MAE
        calc_mae(x = Atest$claimSizeS,
                 y = predict(Afraud.rf,Atest))
        
```
Compare results table
```{r}
ResultsTable = data.frame(c("NoFraud","Fraud","All","NoFraud","Fraud","All","NoFraud","Fraud","All","NoFraud","Fraud","All"))
ResultsTable$Model = c("Logit","Logit","Logit","Rf","Rf","Rf","Logit","Logit","Logit","Rf","Rf","Rf")
ResultsTable$Metric = c("RMSE","RMSE","RMSE","RMSE","RMSE","RMSE","MAE","MAE","MAE","MAE","MAE","MAE")
ResultsTable$Values = c(0.59,0.51,0.60,0.59,0.51,0.60,0.0018,0.083,-0.0014,0.0013,0.092,-0.018)
colnames(ResultsTable) = c("Subset","Model","Metric","Values")
```
